Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 PandaEvolution.tex
Iteration on blackfill: slots + total availability

When the backfill optimization is enabled, users can interrogate Moab about the
number of worker nodes and walltime that would be available to a low-priority job at
that moment in time. If a job is immediately submitted to Titan with that number
of worker nodes and walltime, chances are that Moab will immediately schedule
it, reducing its queue time to a minimum. In this paper, we call this number of
worker nodes and walltime `backfill availability'.

@@

Compared to pilots, backfill has the disadvantage of limiting the amount of
worker nodes that can be requested. Pilots are normal jobs: they can request as
many worker nodes for as much time as a queue can offer. On the contrary, jobs
sized on the basis of backfill availability depend on the number of worker nodes
that cannot be given to any other job in the Titan's queue at that moment in
time.

@@

At any point in time, backfill availability is typically a small fraction of the
total capacity of a resource. Notwithstanding, given the size of Titan this
translates into a substantial capacity. Every year, about 10\% of Titan's
capacity remains unused~\cite{barker2016us}, corresponding to an average of
30,000 unused cores (excluding GPU cores). This equals approximately 270M core
hours per year, roughly 10\% of the overall capacity of WLCG.

@@

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulation in 9 steps. PanDA Broker queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
pulls jobs' input files from the ATLAS DDM service to the OLCF Lustre file
system (Fig.~\ref{fig:panda_broker}:3). PanDA Broker queries Titan's Moab
scheduler about current backfill availability (Fig.~\ref{fig:panda_broker}:2)
and creates an MPI script, wrapping enough ATLAS jobs' payload to fit backfill
availability. PanDA Broker submits the MPI script to the Titan's Torque batch
system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).

>>>>

When the backfill optimization is enabled, users can interrogate Moab about the
number of worker nodes and walltime that would be available to a low-priority
job at that moment in time. If a job is immediately submitted to Titan with that
number of worker nodes and walltime, chances are that Moab will immediately
schedule it, reducing its queue time to a minimum. In this paper, we call this
number of worker nodes and walltime an available `backfill slot'.

@@

Compared to pilots, backfill has the disadvantage of limiting the amount of
worker nodes that can be requested. Pilots are normal jobs: they can request as
many worker nodes for as much time as a queue can offer. On the contrary, jobs
sized on the basis of an available backfill slot depend on the number of worker
nodes that cannot be given to any other job in the Titan's queue at that moment
in time.

@@

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given the
size of Titan this translates into a substantial capacity. Every year, about
10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
average of 30,000 unused cores (excluding GPU cores). This equals
% approximately 270M core-hours per year,
to roughly 10\% of the overall capacity of WLCG.

@@

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulation in 9 steps. PanDA Broker queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
pulls jobs' input files from the ATLAS DDM service to the OLCF Lustre file
system (Fig.~\ref{fig:panda_broker}:3). PanDA Broker queries Titan's Moab
scheduler about the current available backfill slot
(Fig.~\ref{fig:panda_broker}:2) and creates an MPI script, wrapping enough ATLAS
jobs' payload to fit the backfill slot. PanDA Broker submits the MPI script to
the Titan's Torque batch system via RADICAL-SAGA
(Fig.~\ref{fig:panda_broker}:4).

### 2017-04-03 overall.tex
Space saving figures
Iteration conclusions
Iteration on blackfill: slots + total availability

% -----------------------------------------------------------------------------
\subsection{Characterizing the PanDA Broker on Titan}
\label{ssec:panda_titan}

@@

We calculate the total amount of backfill availability over a period of time by:
(i) polling backfill availability at regular intervals during that time window;
(ii) converting the number of worker nodes available and their walltime into
core-hours; (iii) summing the number of core-hours. We call this number of
core-hours `total backfill availability'.

@@

% -----------------------------------------------------------------------------
\subsection{Characterizing the Detector Simulation on Titan}
\label{ssec:panda_titan}

@@

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
    \caption{Comparison between distributions of the time taken by a Geant4
    detector simulation to simulate one event when placing 2 simulations (h1) or
    1 simulation (h2) per CPU. 2 simulation use 16 cores per node, 1 simulation
    8, 1 per compute unit. 50 Events; 1 Titan worker nodes; 16 work threads per
    node; 100 events per node.}
\label{fig:comparison-8-16cores}
\end{figure}

>>>>

% -----------------------------------------------------------------------------
\subsection{Characterizing the PanDA Broker on Titan}
\label{ssec:broker_titan}

@@

We calculate the total amount of backfill availability over a period of time by:
(i) polling the available backfill slots at regular intervals during that time
window; (ii) converting the number of worker nodes available and their walltime
into core-hours; (iii) summing the number of core-hours. We call this number of
core-hours `total backfill availability'.

@@

% -----------------------------------------------------------------------------
\subsection{Characterizing the Detector Simulation on Titan}
\label{ssec:athenamp_titan}

@@

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison_vsquashed.pdf}
    \caption{Comparison between distributions of the time taken by a Geant4
    detector simulation to simulate one event when placing 2 simulations (h1) or
    1 simulation (h2) per CPU. 2 simulation use 16 cores per node, 1 simulation
    8, 1 per compute unit. 50 Events; 1 Titan worker nodes; 16 work threads per
    node; 100 events per node.}
\label{fig:comparison-8-16cores}
\end{figure}
